---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ijtiff

[![Travis-CI Build Status](https://travis-ci.org/rorynolan/ijtiff.svg?branch=master)](https://travis-ci.org/rorynolan/ijtiff)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rorynolan/ijtiff?branch=master&svg=true)](https://ci.appveyor.com/project/rorynolan/ijtiff)
[![codecov](https://codecov.io/gh/rorynolan/ijtiff/branch/master/graph/badge.svg)](https://codecov.io/gh/rorynolan/ijtiff)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ijtiff)](https://cran.r-project.org/package=ijtiff)
![RStudio CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/ijtiff)
![RStudio CRAN monthly downloads](http://cranlogs.r-pkg.org/badges/ijtiff)
[![Rdocumentation](http://www.rdocumentation.org/badges/version/ijtiff)](http://www.rdocumentation.org/packages/ijtiff)
![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)
[![DOI](https://zenodo.org/badge/111798542.svg)](https://zenodo.org/badge/latestdoi/111798542)

## TL;DR

_ImageJ_ sometimes writes channel information in TIFF files in a peculiar way, meaning that most ordinary TIFF-reading softwares don't read this channel information correctly. `ijtiff` knows about _ImageJ_'s peculiarities, so it can be relied upon to read _ImageJ_-written TIFF files correctly.

## Introduction

The _ImageJ_ software (https://imagej.nih.gov/ij) is a widely-used image viewing and processing software, particularly popular in microscopy and life sciences. It supports the TIFF image format (and many others). It reads TIFF files perfectly, however it can sometimes write them in a peculiar way, meaning that when other softwares try to read TIFF files written by _ImageJ_, mistakes can be made. 

The goal of the `ijtiff` R package is to correctly import TIFF files that were saved from _ImageJ_ and to write TIFF files than can be correctly read by _ImageJ_. It may also satisfy some non-_ImageJ_ TIFF requirements that you might have. This is not an extension of the original `tiff` package; it behaves differently. Hence, if this package isn't satisfying your TIFF needs, it's definitely worth checking out the original `tiff` package.

#### The Peculiarity of _ImageJ_ TIFF files

_Note_: If you don't care about the particulars of TIFF files or how this package works on the inside, feel free to skip this subsection.

It is common to use `TIFFTAG_SAMPLESPERPIXEL` to record the number of channels in a TIFF image, however _ImageJ_ sometimes leaves `TIFFTAG_SAMPLESPERPIXEL` with a value of 1 and instead encodes the number of channels in `TIFFTAG_IMAGEDESCRIPTION` which might look something like `"ImageJ=1.51 images=16 channels=2 slices=8"`.

A conventional TIFF reader would miss this channel information (becaus it is in an unusual place). `ijtiff` does not miss it. We'll see an example below. First, let's install the package.


## Installation

### `libtiff`

`ijtiff` requires you to have the `libtiff` C library installed. To install `libtiff`:

* On __Debian Linux__, try `sudo apt-get install libtiff5`, or if that fails, try `sudo apt-get install libtiff4`.
* On __Fedora Linux__, try `sudo yum install libtiff5`, or if that doesn't work, try `sudo yum install libtiff4`.
* On __Mac__, you need [Homebrew](https://brew.sh/). Then in the terminal, run `brew install libtiff`.
* On __Windows__, for most people, no setup is required `r emo::ji("smile")`, but if you experience problems, check out http://gnuwin32.sourceforge.net/packages/tiff.htm.

### Installing the release version of the `ijtiff` R package

You can install `ijtiff` from CRAN (recommended) with:
```{r CRAN installation, eval=FALSE}
install.packages("ijtiff")
```

### Installing the release version of the `ijtiff` R package

You can install the development version from GitHub with:
```{r GitHub installation, eval=FALSE}
if (!require(devtools)) install.packages("devtools")
devtools::install_github("rorynolan/ijtiff")
```
## Reading _ImageJ_ TIFF files

```{r 2 channel path}
path_2ch_ij <- system.file("img", "Rlogo-banana-red_green.tif", 
                           package = "ijtiff")
```
`path_2ch_ij` is the path to a TIFF file which was made in _ImageJ_ from the R logo dancing banana GIF used in the README of Jeroen Ooms' `magick` package. The TIFF is a time-stack containing only the red and green channels of the first, third and fifth frames of the original GIF. Here's the full gif:

![](https://github.com/rorynolan/ijtiff/blob/master/inst/img/Rlogo-banana.tif)

Here are the red and green channels of the first, third and fifth frames of the TIFF:

```{r red and green banana, echo=FALSE, message=FALSE}
library(magrittr)
rgbanana_tif <- system.file("img", "Rlogo-banana-red_green.tif",
                            package = "ijtiff") %>% 
  ijtiff::read_tif()
d <- dim(rgbanana_tif)
rgbanana_reds <- purrr::map(seq_len(d[4]), ~ rgbanana_tif[, , 1, .]) %>% 
  purrr::reduce(rbind)
rgbanana_greens <- purrr::map(seq_len(d[4]), ~ rgbanana_tif[, , 2, .]) %>% 
  purrr::reduce(rbind)
mx <- 2 ^ ceiling(log2(max(rgbanana_tif))) - 1
rgbanana_reds[] %<>% grDevices::rgb(., 0, 0, maxColorValue = mx)
rgbanana_greens[] %<>% grDevices::rgb(0, ., 0, maxColorValue = mx)
plot(as.raster(cbind(rgbanana_reds, rgbanana_greens)))
```

### The original `tiff` package

When we import it with the original `tiff` package:
```{r original tiff import}
img <- tiff::readTIFF(path_2ch_ij, all = TRUE)
str(img)  # 10 images
img[[1]][100:110, 50:60]  # print a section of the first image in the series
```

* We just get a list of `r length(img)` frames, with no information about the `r dim(img[[1]][3])` channels.
* We get annoying warnings about ImageJ's private TIFF tags 50838 and 50839, which are of no interest to the `R` user. 
* The numbers in the image array(s) are (by default) normalized to the range [0, 1].

### The `ijtiff` package

When we import the same image with the `ijtiff` package:
```{r ijtiff import}
img <- ijtiff::read_tif(path_2ch_ij)
dim(img)  # 2 channels, 5 frames
img[100:110, 50:60, 1, 1]  # print a section of the first channel, first frame
```

* We see the image nicely represented as an array of `r dim(img[[1]][3])` channels of `r dim(img[[1]][4])` frames.
* We get no needless warnings.
* The numbers in the image are integers, the same as would be seen if one opened the image with ImageJ.

#### Note
The original `tiff` package reads several types of TIFFs correctly, including many that are saved from _ImageJ_. This is just an example of a TIFF type that it doesn't perform so well with.

## Floating point TIFFs
The original `tiff` package could read but not write floating point (real-numbered) TIFF files. The `ijtiff` package can do both. It automatically decides which type is appropriate when writing.

## Advice for all _ImageJ_ users
Base _ImageJ_ (similar to the `tiff` R package) does not properly open some perfectly good TIFF files^[I think native _ImageJ_ only likes 1, 3 and 4-channel images and complains about the rest, but I'm not sure about this.] (including some TIFF files written by the `tiff` and `ijtiff` R packages).  Instead it gives you the error message: _imagej can only open 8 and 16 bit/channel images_. These images in fact can be opened in _ImageJ_ using the wonderful _BioFormats_ plugin. See https://imagej.net/Bio-Formats.

# Acknowledgement
This package uses a lot of code from the original `tiff` package by Simon Urbanek.

# Contribution
Contributions to this package are welcome. The preferred method of contribution is through a github pull request. Feel free to contact me by creating an issue. Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.