---
title: "ijtiff"
author: "Rory Nolan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ijtiff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height = 6
)
original_files <- dir()
```

## Reading ImageJ TIFF files

```{r 2 channel path}
path_2ch_ij <- system.file("img", "2ch_ij.tif", package = "ijtiff")
```
`path_2ch_ij` is the path to a 2-channel, five-frame image which was saved from ImageJ. 

### The original `tiff` library

When we import it with the original `tiff` library:
```{r original tiff import}
img <- tiff::readTIFF(path_2ch_ij, all = TRUE)
str(img)  # 10 images
img[[1]][100:110, 101:105]  # print a section of the first image in the series
```

* We just get 10 frames, with no information about the two channels.
* We get annoying warnings about ImageJ's private TIFF tags 50838 and 50839, which are of no interest to the `R` user. 
* The numbers in the image array(s) are (by default) normalized to the range [0, 1].

### The `ijtiff` library

When we import the same image with the `ijtiff` library:
```{r ijtiff import}
img <- ijtiff::read_tif(path_2ch_ij)
dim(img)  # 2 channels, 5 frames
img[100:110, 101:105, 1, 1]  # print a section of the first channel, first frame
```

* We see the image nicely divided into 2 channels of 5 frames.
* We get no needless warnings.
* The numbers in the image are integers, the same as would be seen if one opened the image with ImageJ.

`ijtiff` also includes a basic image display function:
```{r display}
ijtiff::display(img[, , 1, 1])  # first channel, first frame
```

#### Note
`tiff` reads several types of TIFFs correctly, including many that are saved from ImageJ. This is just an example of a TIFF type that it doesn't perform so well with.

## Floating point TIFFs
The original `tiff` library could read but not write floating point (real-numbered) TIFF files. The `ijtiff` library can do both. It automatically decides which type is appropriate when writing.

## Advice for all ImageJ users
Base ImageJ (similar to the `tiff` R package) does not properly open some perfectly TIFF files^[I think native ImageJ only likes 1, 3 and 4-channel images and complains about the rest, but I'm not sure about this.] (including some TIFF files written by the `tiff` and `ijtiff` R packages).  Instead it gives you the error message: _imagej can only open 8 and 16 bit/channel images_. These images in fact can be opened in ImageJ using the wonderful _BioFormats_ plugin. See https://imagej.net/Bio-Formats.

## Text Images
TIFF files are limited in which numbers they can represent (they can't go outside the 32-bit range). Real-numbered TIFFs can also lack precision, having only the precision of a 32-bit floating point number. If TIFF isn't good enough, you can use text images.
```{r text images, error=TRUE}
library(ijtiff)
img[1] <- 2 ^ 99  # too high for TIFF
write_tif(img, "img")  # errors
write_txt_img(img, "img")  # no problem
```

```{r cleanup, include = FALSE}
new_files <- setdiff(dir(), original_files)
file.remove(new_files)
```