---
title: "Text Images"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Text Images}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Understanding Text Images

Text images are arrays of numbers stored in a tab-delimited file (https://en.wikipedia.org/wiki/Tab-separated_values), where each location in the file represents a pixel and the value stored there is the pixel intensity.

### Why Use Text Images?

Text images serve two main purposes:

1. **Compatibility**: Many software packages (e.g., Microsoft Excel) don't support saving arrays as TIFF files but can save them as tab-separated text files.

2. **Value Range**: A 32-bit TIFF file can only hold values up to $2^{32} - 1$ (approximately $4 \times 10^9$). Text images have no such limitation, making them useful for storing very large numerical values.

### Limitations

While text images can be useful in specific situations, they have some drawbacks:
* Larger file sizes compared to binary formats
* Slower to read and write
* Less standardized than TIFF format
* Not supported by most image viewing software

## Working with Text Images

### Reading Text Images

To read a text image, use the `read_txt_img()` function:

```{r read-txt-img}
library(ijtiff)
path_txt_img <- system.file("img", "Rlogo-grey.txt", package = "ijtiff")
txt_img <- read_txt_img(path_txt_img)
print(dim(txt_img))  # Show image dimensions
```

Note that `read_txt_img()` expects a tab-separated file (TSV). This is the format that ImageJ uses when saving text images. Other formats like CSV are not supported.

### Writing Text Images

Writing a text image is straightforward using `write_txt_img()`:

```{r write-txt-img}
# Create a simple test image
test_img <- array(1:16, dim = c(4, 4, 1, 1))

# Write it as a text image
out_path <- tempfile(pattern = "txtimg", fileext = ".txt")
write_txt_img(test_img, path = out_path)

# Read it back to verify
read_back <- read_txt_img(out_path)
all.equal(test_img, read_back)
```

### Converting Between TIFF and Text Images

You can convert between TIFF and text image formats using a combination of `read_tif()`, `write_tif()`, `read_txt_img()`, and `write_txt_img()`:

```{r convert}
# TIFF to text
tiff_path <- system.file("img", "Rlogo-grey.tif", package = "ijtiff")
img <- read_tif(tiff_path)
txt_path <- tempfile(fileext = ".txt")
write_txt_img(img, txt_path)

# Text to TIFF
txt_img <- read_txt_img(txt_path)
tiff_path2 <- tempfile(fileext = ".tif")
write_tif(txt_img, tiff_path2)
```

## See Also

* For general image reading and writing, see `vignette("reading-and-writing-images", package = "ijtiff")`
* For working with TIFF tags, see `vignette("tiff-tags", package = "ijtiff")`
* For ImageJ compatibility, see `vignette("the-imagej-problem", package = "ijtiff")`
