---
title: "Working with TIFF Tags"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Working with TIFF Tags}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  crop = TRUE
)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
```

## Quick Start

```{r quickstart}
library(ijtiff)

# Read tags from a TIFF file
tags <- read_tags(system.file("img", "Rlogo.tif", package = "ijtiff"))

# Write a TIFF with custom tags
img <- array(1:4, dim = c(2, 2, 1, 1))
write_tif(img, tempfile(fileext = ".tif"), 
          description = "Test image",
          resolution = c(300, 300))
```

## Understanding TIFF Tags

TIFF (Tagged Image File Format) files contain not just image data, but also metadata in the form of tags. These tags provide important information about the image, such as:

* Image dimensions (width, height)
* Color space and channel information
* Resolution and physical dimensions
* Compression method
* Custom metadata

The `ijtiff` package provides robust support for reading and writing several important TIFF tags.

## Reading TIFF Tags

### Basic Tag Reading

You can read tags from a TIFF file without loading the entire image using `read_tags()`:

```{r read-tags}
path_example <- system.file("img", "Rlogo-banana.tif", package = "ijtiff")
tags <- read_tags(path_example)
str(tags$frame1)  # Show just the first frame's tags
```

### Tag Structure

The tags are returned as a list with one element per frame. Each frame's tags include:

* Image dimensions (ImageLength, ImageWidth)
* Bits per sample
* Samples per pixel (channels)
* Resolution information (XResolution, YResolution, ResolutionUnit)
* Compression method
* And more

## Writing TIFF Tags

### Image Description

Add a description to your TIFF file using the `description` parameter:

```{r write-description}
# Create a simple test image
img <- array(1:4, dim = c(2, 2, 1, 1))
tmp_file <- tempfile(fileext = ".tif")

# Write with description
write_tif(img, tmp_file, description = "My test image", msg = FALSE)

# Read back the description
tags <- read_tags(tmp_file)
cat("Image description:", tags$frame1$description, "\n")
```

### Color Space

Control the photometric interpretation of the image:

```{r write-color-space}
# Write a grayscale image with min-is-black interpretation
img_gray <- array(1:4, dim = c(2, 2, 1, 1))
tmp_file <- tempfile(fileext = ".tif")
write_tif(img_gray, tmp_file, 
          color_space = "min-is-black",  # Default for grayscale
          msg = FALSE)

# Write an RGB image
img_rgb <- array(1:24, dim = c(2, 2, 3, 1))
tmp_file2 <- tempfile(fileext = ".tif")
write_tif(img_rgb, tmp_file2, 
          color_space = "rgb",  # Required for RGB images
          msg = FALSE)

# Read back the color space information
tags_gray <- read_tags(tmp_file)
tags_rgb <- read_tags(tmp_file2)
cat("Grayscale color space:", tags_gray$frame1$color_space, "\n")
cat("RGB color space:", tags_rgb$frame1$color_space, "\n")
```

### Resolution Information

Set image resolution in different units:

```{r write-resolution}
# Write with resolution in inches (default)
write_tif(img, tmp_file, 
          resolution = c(300, 300),  # x and y resolution
          resolution_unit = "inch",
          overwrite = TRUE,
          msg = FALSE)

# Read back the resolution information
tags <- read_tags(tmp_file)
cat("Resolution:", tags$frame1$x_resolution, "x", tags$frame1$y_resolution, 
    "pixels per", tags$frame1$resolution_unit, "\n")
```

### Supported Tags

The `ijtiff` package supports reading and writing several TIFF tags:

* **ImageWidth (256)**: Width of the image in pixels
* **ImageLength (257)**: Height of the image in pixels
* **BitsPerSample (258)**: Number of bits per component
* **Compression (259)**: Compression scheme used
* **PhotometricInterpretation (262)**: Color space of the image data
* **ImageDescription (270)**: ASCII string describing the image
* **SamplesPerPixel (277)**: Number of components per pixel
* **XResolution (282)**: Pixels per ResolutionUnit in the ImageWidth direction
* **YResolution (283)**: Pixels per ResolutionUnit in the ImageLength direction
* **ResolutionUnit (296)**: Unit of measurement for XResolution and YResolution
* **SampleFormat (339)**: Data sample format

## Reading Images with Tags

When you read an image using `read_tif()`, the tags are automatically attached as attributes:

```{r read-with-tags}
# Read an image
img <- read_tif(path_example, msg = FALSE)

# View key attributes (tags)
str(attributes(img)[c("width", "length", "bits_per_sample", "samples_per_pixel")])
```

## TIFF Tags Reference

For a complete reference of all supported TIFF tags and their properties, use `tif_tags_reference()`:

```{r tags-reference}
# View the complete TIFF tags reference
ref <- tif_tags_reference()
head(ref, 20)  # Show first few tags
```

This reference includes information about each tag's:

* Decimal and hexadecimal codes
* Name and short description
* Tag type (baseline or extended)
* C type in libtiff
* Default value
* And more

## Tips for Working with Tags

1. Always check the image description and resolution when working with scientific images
2. When writing images for ImageJ use, set appropriate resolution units
3. The `sample_format` tag is automatically set based on your data type
4. Some tags are read-only and set automatically by the package

## See Also

* For general image reading and writing, see `vignette("reading-and-writing-images", package = "ijtiff")`
* For ImageJ compatibility, see `vignette("the-imagej-problem", package = "ijtiff")`
* For text image support, see `vignette("text-images", package = "ijtiff")`

## Further Reading

* [Library of Congress documentation on TIFF tags](https://www.loc.gov/preservation/digital/formats/content/tiff_tags.shtml)
* [ImageJ TIFF format description](https://imagej.net/ij/docs/guide/146-7.html)
